The \project{unRAVE} catalog

### 2016-09-02 article/unrave.tex
Describe the three-label model training and LOOCV

\usepackage{amsmath}
\usepackage{bm}

% This section can be removed at submission.
% ------------------------------------------
\usepackage{color}
\usepackage{datenumber}

@@

The model we have introduced is generic in the sense that there is no description for how
complex the vectorizing function $\vecv(\ell)$ should be, or how many $K$ labels are used.
Given previous successes using a second-order polynomial model 
\citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}, we will first adopt a three-label ($\teff$, $\logg$, [Fe/H]) model for all (giant and main-sequence) stars. Thus, for our first set of
model experiments:
\begin{eqnarray}\label{eq:vectorizer-three-label}
\vecv(\ell_n) & \rightarrow & \left[1, T_{{\rm eff},n}, \logg_n, [{\rm Fe/H}]_n, T_{{\rm eff},n}^2, \logg_n\cdot{}T_{{\rm eff},n}, \feh_n\cdot{}T_{{\rm eff},n}, \logg_n^2, \feh_n\cdot\logg_n, \feh_n^2\right]
\quad ,
\end{eqnarray}

\noindent{}such that $\vecv(\ell)$ produces the design matrix:
\begin{tiny}
\begin{equation}\label{eq:vectorizer-three-label-design-matrix}
\vecv(\ell) \rightarrow \begin{bmatrix}
	1 & T_{{\rm eff},0} & \logg_{0} & [{\rm Fe/H}]_{0} & T_{{\rm eff},0}^2 & \logg_{0}\cdot{}T_{{\rm eff},0} & [{\rm Fe/H}]_{0}\cdot{}T_{{\rm eff},0} & \logg_{0}^2 & [{\rm Fe/H}]_{0}\cdot{}\logg_1 & [{\rm Fe/H}]_1^2 \\
	\vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots \\
	1 & T_{{\rm eff},n} & \logg_{n} & [{\rm Fe/H}]_{n} & T_{{\rm eff},n}^2 & \logg_{n}\cdot{}T_{{\rm eff},n} & [{\rm Fe/H}]_{n}\cdot{}T_{{\rm eff},n} & \logg_{n}^2 & [{\rm Fe/H}]_{n}\cdot{}\logg_n & [{\rm Fe/H}]_n^2 \\
	\vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots \\
	1 & T_{{\rm eff},N-1} & \logg_{N-1} & [{\rm Fe/H}]_{N-1} & T_{{\rm eff},N-1}^2 & \logg_{N-1}\cdot{}T_{{\rm eff},N-1} & [{\rm Fe/H}]_{N-1}\cdot{}T_{{\rm eff},N-1} & \logg_{N-1}^2 & [{\rm Fe/H}]_{N-1}\cdot{}\logg_{N-1} & [{\rm Fe/H}]_{N-1}^2
\end{bmatrix}\quad .
\end{equation}
\end{tiny}

\subsubsection{A three-label model for main-sequence and giant stars}\label{sec:a-simple-joint-model}

\subsubsection{A three-label model for main-sequence stars}\label{sec:a-simple-model-ms-stars}

% worked but didn't show the structure in detail.
% Split into two different models with partial overlap.
% One for giants and one for dwarfs.
% The labels for stars in the training sets are shown in Figure X.

% We trained all three models.
% Note the differences in metallicity scales.
% Look at spectral derivatives.
% Show the model internals are relatable.
% Map spectral derivatives from giant model to main-sequence model.
% Update the model with those new spectral derivatives.

% How we combined the results into one.
% Refer to test set HRD.

@@

% TODO: Replace ZENODO-URL with a data url
\noindent{}Source code for this project is available at \texttt{\giturl}.
This document was compiled from revision hash \texttt{\githash} in that repository.
Inferred labels and covariance matrices are available electronically with 
associated metadata at \texttt{\url{ZENODO-URL}} \citep{DATA_REPOSITORY}.  
Please note that it is a condition of using these results that the fifth \rave\ 
data release by \citet{Kunder_2016} must (also) be cited, as the work presented 
here would not have been possible without the tireless efforts of the entire 
\rave\ collaboration, past and present.

\acknowledgements
A.~R.~C. warmly thanks Jonathan Bird (Vanderbilt), Sergey Koposov (Cambridge),
and Sven Buder (MPIA).
This research made use of: Astropy, a community-developed core Python package for
Astronomy \citep{astropy}, NASA's Astrophysics Data System Bibliographic Services;
and \project{TOPCAT} \citep{Taylor_2005}.
This work was partly supported by the European Union FP7 programme through ERC 
grant number 320360. Funding for KH has been provided through the Simons Foundation Society of Fellows and the Marshall Scholarship.
Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
University; the Australian Research Council; the French National Research Agency;
the German Research Foundation (SPP 1177 and SFB 881); the European Research 
Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
Padova; The Johns Hopkins University; the National Science Foundation of the USA
(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
Research Council of Canada; the Slovenian Research Agency; the Swiss National 
Science Foundation; the Science \& Technology Facilities Council of the UK; 
Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
and Sydney. The RAVE web site is https://www.rave-survey.org.

>>>>

\usepackage{amsmath}
\usepackage{bm}
\usepackage{xcolor}

\definecolor{unoffensive-warning}{HTML}{B4DCED}

% This section can be removed at submission.
% ------------------------------------------
%\usepackage{color}
\usepackage{datenumber}

@@

The model we have introduced is generic in the sense that there is no description for how
complex the vectorizing function $\vecv(\ell)$ should be, or how many $K$ labels are used.
Given previous successes using a second-order polynomial model 
\citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}, we will first adopt a three-label ($\teff$, $\logg$, [Fe/H]) model for all (giant and main-sequence) stars. Thus, for our first set of
model experiments,
\begin{eqnarray}\label{eq:vectorizer-three-label}
\vecv(\ell_n) & \rightarrow & \left[1, T_{{\rm eff},n}, \logg_n, [{\rm Fe/H}]_n, T_{{\rm eff},n}^2, \logg_n\cdot{}T_{{\rm eff},n}, \feh_n\cdot{}T_{{\rm eff},n}, \logg_n^2, \feh_n\cdot\logg_n, \feh_n^2\right]
\quad ,
\end{eqnarray}

\noindent{}such that $\vecv(\ell)$ produces the design matrix:
\begin{eqnarray}
	\vecv(\ell) \rightarrow \begin{bmatrix} \vecv(\ell_0) \\ \vdots \\ \vecv(\ell_{N-1}) \end{bmatrix} \quad .
\end{eqnarray}

\stub{We trained the model initially with no regularization. Could reproduce general structure of a main-sequence and HRD, but the intricacies of the structure was gone. Model appeared to be too simple. However we can use this model to identify dwarfs/giants,etc. We may revisit this model later}

\subsubsection{A 3-label model for the main-sequence}
\label{sec:the-ms-model}

We constructed a three-label quadratic model for the main-sequence and sub-giant
stars in the \epic\ training set.  In order to set the regularization hyperparameter
$\Lambda$ for this model, we trained 30 models with different regularization strengths,
spaced evenly in logarithmic space between $\Lambda = 10^{-3}$ to $\Lambda = 10^{3}$.
We then performed leave-one-out cross-validation for each model.  Specifically, for 
each star in the training set: we removed the star; trained the model; and then 
inferred labels from the removed star as if it was a test object. We also performed 
leave-one-out cross-validation on an unregularized ($\Lambda = 0$) model, which we 
will use as the basis for comparison.  For every model we calculated the bias and
root-mean-square (RMS) deviation between the input labels and those inferred through
cross-validation.

We show the \emph{percentage difference} in the RMS deviation of the labels with respect
to the unregularized model in Figure \ref{fig:set-ms-model-hyperparameter}.  The upper 
and lower envelope represent the boundaries across all labels, showing that the RMS
decreased in \emph{all} labels with increasing regularization.  We found similar
improvements in the bias values, however these were already minimal in the unregularized
case.  The improvement in RMS reaches a minimum value near $\Lambda = X.XX$, where
we achieve RMS deviations that are about 30\% better than the unregularized model.
Based on this improvement we set $\Lambda = X.XX$ for the three-label main-sequence model.
At this regularization strength, the bias and RMS values found by leave-one-out
cross validation are, respectively: $X$~K and $XX$~K for $\teff$, $X.XX$~dex and $X.XX$~dex
for $\logg$, with $X.XX$~dex and $X.XX$~dex for [Fe/H].

We made no \emph{a priori} judgements on whether a star was likely a giant
or a dwarf star; we inferred labels for all \rave\ spectra using this 
main-sequence model.  The resulting labels for the survey sample are shown
in Figure \ref{fig:hrd-from-individual-models}.  Significant structure is 
visible.  The increased density of Solar-type stars is consistent with \rave\
observing stars in the local neighbourhood, and the high number of turnoff
and main-sequence stars relative to the subgiant branch is expected from 
the relative lifetimes of these evolutionary phases.  An overdensity
of stars near the giant branch clump is visible, which is likely due to
giant stars existing in the test set, but not in the training set.  In
later sections we will describe how our results from multiple models were
compiled to produce a single catalog.

\subsubsection{A 3-label model for the main-sequence and giant branch}
\label{sec:the-joint-model}

% Chose the regularization hyperparameter.

@@

% TODO: Replace ZENODO-URL with a data url
\noindent\colorbox{unoffensive-warning}{\parbox{\dimexpr\linewidth-2\fboxsep}{
\noindent{}Source code for this project is available at \texttt{\giturl}\hspace{-0.5em},
and this document was compiled from revision hash \texttt{\githash} in that repository.
Inferred labels and covariance matrices are available electronically with 
associated metadata at \texttt{\url{ZENODO-URL}} \citep{DATA_REPOSITORY}.  
Please note that it is a condition of using these results that the fifth \rave\ 
data release by \citet{Kunder_2016} must (also) be cited, as the work presented 
here would not have been possible without the tireless efforts of the entire 
\rave\ collaboration, past and present.
}}

\acknowledgements
A.~R.~C. warmly thanks Jonathan Bird (Vanderbilt), Sven Buder (MPIA), and
Sergey Koposov (Cambridge).
This research made use of: Astropy, a community-developed core Python package for
Astronomy \citep{astropy}, NASA's Astrophysics Data System Bibliographic Services;
and \project{TOPCAT} \citep{Taylor_2005}.
This work was partly supported by the European Union FP7 programme through ERC 
grant number 320360. Funding for KH has been provided through the Simons Foundation Society of Fellows and the Marshall Scholarship.
Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
University; the Australian Research Council; the French National Research Agency;
the German Research Foundation (SPP 1177 and SFB 881); the European Research 
Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
Padova; The Johns Hopkins University; the National Science Foundation of the USA
(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
Research Council of Canada; the Slovenian Research Agency; the Swiss National 
Science Foundation; the Science \& Technology Facilities Council of the UK; 
Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
and Sydney. The RAVE web site is https://www.rave-survey.org.
